---
title: "Practical Question - PS2"
author: "Pedro Henrique"
date: "26/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
setwd("~/Área de Trabalho/ProblemSet2/")

# Got data from this link https://www.ibge.gov.br/cidades-e-estados/rj.html
rj <- read.csv('https://raw.githubusercontent.com/pedrohcrocha/Econometrics/main/Problem%20Set%202/rj_density.csv') %>% 
  select(-X) %>% 
  mutate(area.pon = as.character(area.pon))

df <- readRDS('census_rj_pessoas_ps2.Rds')
df <- df %>% 
  filter(work.active == 1, age >= 15, age <= 75, total.monthly > 0) %>% 
  mutate(log.monthly.wages = log(total.monthly)) %>% 
  mutate(female = ifelse(gender == 2, 1, 0)) %>% 
  mutate(area.pon = substr(as.character(area.pon), 1,7))

df <- left_join(df, rj, by="area.pon")
df <- df %>% 
  mutate(log_density = log(density))

`Log Monthly Wages` <- df$log.monthly.wages
logearn_m <- df %>% filter(female == 0) %>% select(log.monthly.wages) %>% pull()
Nm = length(logearn_m)
logearn_f <- df %>% filter(female == 1) %>% select(log.monthly.wages) %>% pull()
Nf = length(logearn_f)
s <- sd(df$`Log Monthly Wages`)
s_m <- sd(logearn_m)
s_f <- sd(logearn_f)

```

# Question 6

Let's use a uniform kernel to build a kernel density estimator function:

```{r uniform kernel density}

# Uniform kernel function
uniform_kernel_function <- function(z, delta = 1.3510){
  K <- 0.5 * as.numeric(abs(z)<1)
  return(K)
}
```

Given the sample code showed in class, for this question, it's better not to reinvent the wheel. Therefore, we adapt the sample code. We end up with:

```{r kernel estimation}
# Kernel density estimator for each sex
kernel_density_estimator_male <- function(x){
  z <- (logearn_m - x)/h
  K <- uniform_kernel_function(z)
  s <- sum(K)
  f_hat <- s/(Nm*h)
  return(f_hat)
}

kernel_density_estimator_female <- function(x){
  z <- (logearn_f - x)/h
  K <- uniform_kernel_function(z)
  s <- sum(K)
  f_hat <- s/(Nf*h)
  return(f_hat)
}

# We can construct a vector of values to catch various data points of our variable
x = seq(min(df$log.monthly.wages), max(df$log.monthly.wages),0.1)

# Log monthly earnings for males and sample size
logearn_m <- df %>% filter(female == 0) %>% select(log.monthly.wages) %>% pull()
Nm = length(logearn_m)

# Log monthly earnings for females and sample size
logearn_f <- df %>% filter(female == 1) %>% select(log.monthly.wages) %>% pull()
Nf = length(logearn_f)

# Estimated kernel values
h = 0.545
yf <- sapply(x, kernel_density_estimator_female)
ym <- sapply(x, kernel_density_estimator_male)

# The plot
`Log Monthly Wages` <- df$log.monthly.wages
hist(`Log Monthly Wages`, freq=FALSE)
lines(x, yf, col = "red")
lines(x, ym, col = "blue")
legend("topright", legend=c("Female", "Male"), col=c("red", "blue"), lty=1, cex=0.8)
legend("topleft", legend=c("h=", h))
```

# Question 7

The optimum bandwidth is given by:

$$ h^* = \delta \left( \int f''(x_0)^2 \,dx_0) \right)^{0.2} N^{-0.2} $$

Assuming that our data is normally distributed, then:

$$ \int f''(x_0)^2 \,dx_0 = \frac{3}{8\sqrt{\pi}\sigma^5} = \frac{0.2116}{\sigma^5} $$

Therefore, the optimal bandwidth is:
$$ h^*  = \delta \left(\frac{0.2116}{\sigma^5}\right)^{0.2} N^{-0.2} = 1.3643\delta N^{-0.2}s $$  

The sample standard deviation for the general data and for each sex are:

``` {r sd}
s <- sd(df$log.monthly.wages)
s

s_m <- sd(logearn_m)
s_m

s_f <- sd(logearn_f)
s_f
```


From table 9.1 of Cameron et Privadi (2005), we have the value of $\delta$. Hence, the value of the optimal bandwidths for male and female, respectivaly, are:


``` {r optimal h}
h_star_m <- 1.3643*1.3510*(Nm^-0.2)*s_m
h_star_m

h_star_f <- 1.3643*1.3510*(Nf^-0.2)*s_f
h_star_f
```

Recalculating the estimated results, we end up with:

```{r optimal band result, echo=F}
# Estimated kernel values using the optimum bandwidth
h = h_star_f
yf_star <- sapply(x, kernel_density_estimator_female)

h = h_star_m
ym_star <- sapply(x, kernel_density_estimator_male)

# The plot
`Log Monthly Wages` <- df$log.monthly.wages
hist(`Log Monthly Wages`, freq=FALSE, ylim = c(0,1))
lines(x, yf_star, col = "red")
lines(x, ym_star, col = "blue")
legend("topright", legend=c("Female", "Male"), col=c("red", "blue"), lty=1, cex=0.8)
legend("topleft", legend=c("hf*=", round(h_star_f,3), 'hm*=', round(h_star_m,3)), cex=0.8)
```

# Question 8

```{r density}
epanechnikov <- function(z) {0.75*(1-z^2)*as.numeric(abs(z)<1)}  


log_density <- df$log_density
log.monthly.wages <- df$log.monthly.wages

local_constant_estimator <- function(x) {
  z = (logdensity - x)/h
  k <- epanechnikov(z)
  m_hat <- sum(k*log.monthly.wages)/sum(k)
  return(m_hat)
}

x <- seq(min(log_density), max(log_density), 0.1)

h <- 0.1
y_hat <- sapply(x, local_constant_estimator)
plot(x, y_hat)


h <- 0.2
y_hat <- sapply(x, local_constant_estimator)
plot(x, y_hat)


h <- 0.3
y_hat <- sapply(x, local_constant_estimator)
plot(x, y_hat)

```



